A Large-Scale Study of the Use of Eval in R

### 2020-08-29 inst/paper/main.tex
Tidying
remove dots from numnbers

%\input{tag/enrichDataset.tex}
\newcommand{\UndefinedEvalsRnd}{285K\xspace}
\newcommand{\UndefinedEvals}{285,289\xspace}
\newcommand{\PercentUndefinedEval}{3.3\%\xspace}
\newcommand{\UndefinedEvalExpressionsRnd}{262\xspace}
\newcommand{\UndefinedEvalExpressions}{262\xspace}
\newcommand{\UndefinedOutOfCorpusRnd}{0\xspace}
\newcommand{\UndefinedOutOfCorpus}{0\xspace}
\newcommand{\UndefinedOutOfCoprusPercent}{0\%\xspace}
\newcommand{\MaxCallSitesEnrich}{114\xspace}
\newcommand{\OneCallSitesEnrich}{61\xspace}
\newcommand{\NoCallSitesEnrich}{297\xspace}
%\input{tag/patterns.tex}
\newcommand{\PatternLibRnd}{4K\xspace}
\newcommand{\PatternLib}{4,278\xspace}
\newcommand{\PatternLibRatio}{.05\%\xspace}
\newcommand{\PatternLibSourceRnd}{370\xspace}
\newcommand{\PatternLibSource}{370\xspace}
\newcommand{\PatternLibLocalRnd}{100\xspace}
\newcommand{\PatternLibLocal}{100\xspace}
\newcommand{\PatternReadRnd}{59K\xspace}
\newcommand{\PatternRead}{59,776\xspace}
\newcommand{\PatternReadRatio}{.7\%\xspace}
\newcommand{\PatternReadSlotAccessRatio}{88\%\xspace}
\newcommand{\PatternReadDatatableRatio}{27\%\xspace}
\newcommand{\PatternAssignRnd}{68K\xspace}
\newcommand{\PatternAssign}{68,754\xspace}
\newcommand{\PatternAssignRatio}{.9\%\xspace}
\newcommand{\PatternAssignArrowRatio}{18\%\xspace}
\newcommand{\PatternAssignArrowMagrittrRatio}{59\%\xspace}
\newcommand{\PatternAssignSuperarrowRatio}{81\%\xspace}
\newcommand{\PatternAssignSuperarrowPlyrRatio}{99\%\xspace}
\newcommand{\PatternManualPackagesRnd}{20\xspace}
\newcommand{\PatternManualPackages}{20\xspace}
\newcommand{\PatternManualCallsitesRnd}{161\xspace}
\newcommand{\PatternManualCallsites}{161\xspace}

@@

\newcommand{\eval}{\texttt{eval}}
\newcommand{\evals}{{\sf eval}s\xspace}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}
\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

R supports reflection and meta-programming. The function \c{substitute(e,envir)}
yields the parse tree of the expression \c{e} after performing substitutions
defined by the bindings in the \c{envir} parameter.

@@

\noindent
R allows programmatic manipulation of parse trees, which are themselves
first class objects. They are evaluated using the \c{eval()}
function. There are three alternative forms  of \eval in the \c{base} package:
\c{evalq}, \c{eval.parent} and \c{local}.

\eval function in R takes three arguments as shown below.
\begin{lstlisting}
 eval(expr,
      envir = parent.frame(),
      enclos = if(is.list(envir) || is.pairlist(envir))
               parent.frame() else baseenv())
\end{lstlisting}

\c{expr} is the value to be evaluated, \c{envir}
is the environment in which it is to be evaluated, and \c{enclos} is
relevant when \c{envir} is a list or a data frame, then it specifies where R
looks for objects not found in \c{envir}. This can be \c{NULL} (interpreted
as the base package environment, \c{baseenv()}) or an environment.

When called, \eval evaluates \c{expr} in the environment specified by
\c{envir} and returns the computed value. If \c{envir} is missing, then the
default is \c{parent.frame()} (the environment where the call to \eval was
made).  Values to be evaluated can be of types \c{call} (AST for function
calls) or \c{expression} (AST nodes for arbitrary expressions), \c{name}
(when the symbol is looked up in the current scope and its binding is
evaluated), promise, or any of the basic types such as vectors and
environments (which are returned unchanged).

@@

@@

\begin{figure}[!h]\centering\includegraphics[width=\linewidth]
  {corpus.pdf}\caption{CRAN packages}\label{fig:corpus}
\end{figure}

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests},\footnote{We had to exclude most tests because the {testthat}
  harness uses \eval and thus causes the entire test to register as an \eval
  call.} there are also \emph{examples} and \emph{vignettes}. Examples are R
code snippets extracted from package documentation into scripts files.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R
code that can be similarly extracted into R scripts. R provides builtin
functions for tangling the embedded code into files.  The selected packages
contain \CorpusPackagePrograms programs with \CorpusPackageProgramsCodeRnd
lines of R code: \CorpusExamplesProgramsRnd examples with
\CorpusExamplesCodeRnd lines of code, and \CorpusVignettesProgramsRnd
vignettes.

@@

To represent end-user code in the corpus, we turned to Kaggle, an online
platform for data-science and machine-learning. The website allows people to
share data-science competitions, a data-analysis problem together with data,
for which users try to find the best solution. The solutions, called
{kernels}, are then shared back to the platform in the form of either plain
scripts or as notebooks.  One of the most popular competition is about
predicting passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our
corpus.

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats,\footnote{We use {\sf rmarkdown} to convert from notebooks to R.} we
found that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next
to various runtime exceptions, common problems were parse errors and
misspelled package names.  The final set contains \CorpusFinishedKaggle
kernels with \CorpusFinishedKaggleCodeRnd lines of R code.

@@

This section presents an overview of the usage of \eval in our corpus. Richards
et al. reported 82\% of JavaScript-enabled web pages using \eval~\cite{ecoop11}.
The case for R is more striking with 100\% of the programs call the \eval
function.\footnote{We combine calls to {\sf eval}, {\sf eval.parent}, {\sf
    evalq} and {\sf local}.} This high number is explained by the fact that the
R implementation, its core packages, use \eval extensively. We differentiate
three types of eval based on their origin:
%
\begin{compactitem}[$-$]
  \item a {\bf core eval} is one that occurs in the code of a core package,
  \item a {\bf dependent eval} occurs in a package imported by the target
  package, and
  \item an {\bf own eval} is one that occurs in the source code of the target
  package.
\end{compactitem}

@@

%% SORRY  I HAD TO REFORMAT
\begin{table}[!h] \centering
 \begin{tabular}{r|c|c|c} \hline
Source & Core    & Dependent                      & Own \\\hline
CRAN   & 5.6M &  867K    & 259K \\
       & (1 / 337 / 46K)& (0 / 52 / 58K ) & (0 / 15 / 15K) \\\hline
Kaggle &  1.2M &  470K   & 0\\
 &  (0 / 721 / 18K) & (0 / 290 / 20K) & (0 / 0 / 0)\\\hline
\end{tabular}\caption{Eval calls in CRAN and Kaggle}\label{A}\end{table}

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. \OneCallSitesEnrich{} packages have a single call site, and one
package, \emph{metafor}, has \MaxCallSitesEnrich{} call sites of
\eval. \NoCallSitesEnrich{} packages have no calls to \eval. The core packages
have \NbCoreEvalCallSites{} call sites of \eval and Kaggle has none. These
numbers are under-approximations as \eval may be called reflectively,
through aliases, or passed to a higher-order function.

@@

In this section we turn our attention to the expression passed for
evaluation to \eval. The \eval function accepts all types of R objects as
inputs, but in the majority of cases (\AllExpressionInputEvalCallPerc), it
is called with expressions. The remaining \AllValueInputEvalCallPerc are
values, \ie, they evaluate to themselves.
%
Passing a value into an eval is effectively an no-op so we look closely at the
expression types the call sites pass to eval across all the runs.
\AllValueInputEvalSitePerc of all call sites always pass values,
\AllExpressionInputEvalSitePerc pass expressions and a very large proportion,
\AllPolymorphicInputEvalSitePerc, pass both expressions and values as inputs
across all program runs. We inspected the eval callsites that accept only values
or both values and expression as inputs. The top three of these call sites
contribute to over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this function finds
the default values for parameters with those names in the caller function. Since
the default value for a function parameter can be a complex expression or a
benign value such as \c{NULL}, the \c{match.arg} function evaluates it through a
call to \eval. This makes the \eval inside \c{match.arg} ``polymorhpic''. The
second such \eval call originates from \ggplot, a popular R package for making
plots. \ggplot has it's own object system, \ggproto, which uses \eval inside the
\ggproto function to determine the super-class from which it inherits (the
default value is \c{NULL} but in general it could be an arbitrarily expression).
The third \eval call is in core R \c{str} function used to print a summary of R
objects. It internally maps \eval function on the bindings of an environment
which results in inputs of multiples types being passed to \eval.

@@

We looked at the top ten expressions passed to core and package \eval
calls. The most frequent ten expressions to eval calls from core R
contribute to 85\% of all \eval calls.

\begin{table}[!h] \centering
\begin{tabular}{@{}l|rr@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{c("auto", "shell", "radix")} & 198,7105 & 29\%\\
\c{c("auto", "shell", "quick", "radix")} & 1,593,169  & 23\%\\
\c{\{info <- loadingNamespaceInfo(...} & 1,008,632 &       14\%\\
\c{c("onLoad", "attach", "detach", "onUnload")}   & 470,566 &      6.9\%\\
\c{c("append", "prepend", "replace")} &              261,587&       3.9\% \\
\c{c("left", "right", "centre", "none")} & 162,086     & 2.4\%\\
\c{c("no", "ifany", "always")}   &                71,580 &       1.1\%\\
\c{c("pearson", "kendall", "spearman")}  & 72,962 &      1.1\%\\
\c{NULL}& 75330  &      1.1\% \\
\c{Symbol}&                 6,6279&       1\%\\\hline
\end{tabular}
\caption{Top ten eval calls in Core}\label{B}
\end{table}

@@

The most frequent ten expressions to eval calls from CRAN packages
contribute to 77.1\% of all \eval calls.

\begin{table}[!h]  \centering
\begin{tabular}{@{}l@{~}|@{~}r@{~}r@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{Environment} &                                  989302   & 61\%\\
\c{column[rows] <<- what} &                        55677    & 3.5\%\\
\c{function(value) freduce(value, `_function_list`)} & 37251& 2.3\%\\
\c{NULL} &                         32005    & 2\%\\
\c{List} &                         22293    & 1.4\%\\
\c{c("default", "default2012", "default2011" ...}& 20610    & 1.3\%\\
\c{force(..1)}            &                        20461    & 1.3\%\\
\c{alist(`_spec`)}       &                         18532    & 1.2\%\\
\c{inner}               &                          18530    & 1.2\%\\
\c{String Vector}      &                           17487     & 1.1\%\\
\end{tabular}\caption{Top ten eval calls in CRAN} \label{C}
\end{table}

The expression \c{Environment} occurs because of the four callsites
explained above, \c{ggplot2::ggproto}, \c{R6::generator_funs},
\c{future::backtrace} and \c{RModel::str.RMmodel}. The next expression,
\c{column[rows] <<- what}, is used inside the \c{plyr::rbind.fill} function
to merge data frames by assigning concatenated vectors to rows. The \c{<<-}
operator is interesting in that it skips the current scope and assigns in a
parent scope in which the variable is already present. In our corpus, all
these \eval calls contribute to a single side-effect. The expression
\c{function(value) freduce(value, `_function_list`)} arises from the
\c{magrittr::\%>\%} function which is a pipe operator that pipes the output
of previous command to the next one. The expression is evaluated in a custom
environment to create a function binding for evaluating the components of
the pipe.\c{String Vector} and \c{List} also arise from the same function
when a string or a list is piped using the \c{\%>\%} function into the next
expression.  The \c{NULL} arises from \c{R6::generator_funs} function when
the \eval is passed a \c{NULL} argument by the \c{DataMask_generator}
package. The \c{c("default", "default2012", "default2011" ...} pattern
arises from \c{copula::polyG} where it reflectively access the default
expression for its formal parameter and evaluates it. The \c{force(..1)} and
\c{alist(`_spec`)} patterns occur in \c{glue::glue_data} function which
concatenates and interpolates strings. The two patterns occur because the
function captures unevaluated unnamed arguments and maps the evaluation of
\c{force(..1)} on them.  The \c{force} function forces promises and returns
the result of evaluation. The \c{inner} pattern arises from
\c{glue::identity_transformer} which enables the creation of custom
transformation functions for affecting the interpolation and concatenation
of input by the \c{glue} package.

\subsection{Operation Mix}

The JavaScript paper reported on the operation mix within \evals. This means
the proportion of read, write, and call bytecodes performed by the
JavaScript interpreter. They found that some web pages were creating objects
more frequently and writing into their fields.  In R, nearly every operation
performed in the interpreter is a function call. We observe 2.9G calls and,
of those 125M happen inside \eval (there are 463M native calls, of which
only 31M are in \eval).

{\bf there is no good reason for not counting reads and writes, we can...}

%% ...The following is fluff ...

%% Qualitatively, we observe all kinds of operations being executed inside
%% \eval calls, from numerical computing for statistical modeling to reflection
%% and metaprogramming for implementing DSLs. Even testing frameworks such
%% \c{genthat} heavily rely upon \eval to execute test code snippets. The core
%% R implementation is rife with uses of \eval for mundane things like
%% accessing default values of formal parameters and in statistical packages
%% for computation.

\subsection{Provenance}

In JavaScript, the provenance of the string passed into \eval is a matter of
concern as \eval can be used to breach the security of websites. The main
danger comes from string that were obtained from external sources. If they
are allowed to flow into an \eval, then an attacker could hijack the
program.

{\bf what have they found...}

There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}

We observe 5.2\% cases where \eval directly evaluates the output of
\c{substitute}, 0.7\% cases where output of \c{parse} is read directly and
only 704 cases for \c{expression}. Most expressions consumed by \eval are
generated by other functions.

\c{eval(parse(...))} can be used for dynamic code loading. This forms the
core of \c{source} and \c{sys.source} functions in R that are commonly used
for loading code in R files in interactive settings. We investigated the
number of cases in which the output of \c{parse} and its variants is passed
to \eval, directly or transitively by tainting their output. This
corresponds to \PercentParsedCallSites of the total \eval call sites and
\PercentParsedEvals of the eval calls. We observed that very few of the eval
call (\NbParseFilesRnd in total) consume the result of calling \c{parse} on
a file. Most of the eval calls consume the result of calling \c{parse} on a
string.  We also identified one function in core R,
\c{invokeRestartInteractively} that prompts the user for input, parses it,
and passes it to \eval.

@@

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \eval can
access the local scope and the global scope. In R, \eval has an argument
\c{envir} to indicate in which environment its first argument must be evaluated.
By default, it is equal to \c{parent.frame()}, 
which is the parent environment of \eval call. This default argument
happens in \DefaultEnvirExprPercent of the eval calls.

This argument can also be \c{NULL}, a \c{List} or a data frame. This happens
in 3.5\% of the cases: \eval copies the fields of the list or data frame
and creates bindings for them in a new environment. This pattern is used to
evaluate formulas which can directly refer to the fields of the data.  The
\c{envir} argument can also be a number $n$. It means that the environment
in which the expressions is evaluated will be the result of \c{sys.call(n)}
where $n$ refers to the $n$-th stack frame.

@@

Richards et al.~\cite{ecoop11} identified a number of frequently occurring \eval
patterns which were detected using simple regular expressions on the \eval
strings. In the case of R we match the patterns on the recorded AST.

>>>>

%\input{tag/enrichDataset.tex}
\newcommand{\UndefinedEvalsRnd}{285K\xspace}
\newcommand{\UndefinedEvals}{285,289\xspace}
\newcommand{\PercentUndefinedEval}{3.3\%\xspace}
\newcommand{\UndefinedEvalExpressionsRnd}{262\xspace}
\newcommand{\UndefinedEvalExpressions}{262\xspace}
\newcommand{\UndefinedOutOfCorpusRnd}{0\xspace}
\newcommand{\UndefinedOutOfCorpus}{0\xspace}
\newcommand{\UndefinedOutOfCoprusPercent}{0\%\xspace}
\newcommand{\MaxCallSitesEnrich}{114}
\newcommand{\OneCallSitesEnrich}{61}
\newcommand{\NoCallSitesEnrich}{297}
%\input{tag/patterns.tex}
\newcommand{\PatternLibRnd}{4K\xspace}
\newcommand{\PatternLib}{4,278\xspace}
\newcommand{\PatternLibRatio}{.05\%\xspace}
\newcommand{\PatternLibSourceRnd}{370\xspace}
\newcommand{\PatternLibSource}{370\xspace}
\newcommand{\PatternLibLocalRnd}{100\xspace}
\newcommand{\PatternLibLocal}{100\xspace}
\newcommand{\PatternReadRnd}{59K\xspace}
\newcommand{\PatternRead}{59,776\xspace}
\newcommand{\PatternReadRatio}{.7\%\xspace}
\newcommand{\PatternReadSlotAccessRatio}{88\%\xspace}
\newcommand{\PatternReadDatatableRatio}{27\%\xspace}
\newcommand{\PatternAssignRnd}{68K\xspace}
\newcommand{\PatternAssign}{68,754\xspace}
\newcommand{\PatternAssignRatio}{.9\%\xspace}
\newcommand{\PatternAssignArrowRatio}{18\%\xspace}
\newcommand{\PatternAssignArrowMagrittrRatio}{59\%\xspace}
\newcommand{\PatternAssignSuperarrowRatio}{81\%\xspace}
\newcommand{\PatternAssignSuperarrowPlyrRatio}{99\%\xspace}
\newcommand{\PatternManualPackagesRnd}{20\xspace}
\newcommand{\PatternManualPackages}{20\xspace}
\newcommand{\PatternManualCallsitesRnd}{161\xspace}
\newcommand{\PatternManualCallsites}{161\xspace}

@@

\newcommand{\eval}{\c{eval}}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}
\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

R supports reflection and meta-programming. The function \c{substitute(e,envir)}
yields the parse tree of the expression \c{e} after performing substitutions
defined by the bindings in \c{envir}.

@@

\noindent
R allows programmatic manipulation of parse trees, which are themselves
first class objects. They are evaluated using the \c{eval()}
function. There are three alternative forms  of \eval in the \c{base} package:
\c{evalq}, \c{eval.parent}, \c{local}.

\begin{lstlisting}
 eval(expr,
      envir = parent.frame(),
      enclos = if(is.list(envir) || is.pairlist(envir))
               parent.frame() else baseenv())
\end{lstlisting}

Arguments are as follows: \c{expr} is the value to be evaluated, \c{envir}
is the environment in which it is to be evaluated, and \c{enclos} is
relevant when \c{envir} is a list or a data frame, then it specifies where R
looks for objects not found in \c{envir}. This can be \c{NULL} (interpreted
as the base package environment, \c{baseenv()}) or an environment.

When called, \eval evaluates \c{expr} in the environment specified by
\c{envir} and return the computed value. If \c{envir} is missing, then the
default is \c{parent.frame()} (the environment where the call to \eval was
made).  Values to be evaluated can be of types \c{call} (AST for function
calls) or \c{expression} (AST nodes for arbitrary expressions), \c{name}
(when the symbol is looked up in the current scope and its binding is
evaluated), promise, or any of the basic types such as vectors and
environments (which are returned unchanged).

@@

\begin{figure*}[!tb]\centering\includegraphics[width=.8\linewidth]
  {corpus.pdf}\caption{CRAN packages}\label{fig:corpus}
\end{figure*}

@@

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests},\footnote{We had to exclude most tests because the {testthat}
  harness uses \eval and thus causes the entire test to register as an \eval
  call.} there are also \emph{examples} and \emph{vignettes}. Examples are R
code snippets extracted from package documentation into scripts files.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R
code that can be similarly extracted into R scripts. R provides builtin
functions for tangling the embedded code into files.  The selected packages
contain \CorpusPackagePrograms programs with \CorpusPackageProgramsCodeRnd
lines of R code: \CorpusExamplesProgramsRnd examples with
\summaryrnd{CorpusExamplesCode}\footnote{This notation represents $sum$
  ($min$ / $mean$ / $max$)} lines of code, and \CorpusVignettesProgramsRnd
vignettes.%%% with \summaryrnd{CorpusVignettesCode} lines of code.

@@

To represent end-user code in the corpus, we turned to
Kaggle,\footnote{\url{https://www.kaggle.com}} an online platform for
data-science and machine-learning. The website allows people to share
data-science competitions, a data-analysis problem together with data, for
which users try to find the best solution. The solutions, called {kernels},
are then shared back to the platform in the form of either plain scripts or
as notebooks.  One of the most popular competition is about predicting
passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our
corpus.

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats,\footnote{We use rmarkdown package to convert from Jupyer notebook
  to Rmarkdown as well as to convert from Rmarkdown to plain R.} we found
that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next
to various runtime exceptions, common problems were parse errors and
misspelled package names.  The final set contains \CorpusFinishedKaggle
kernels with \summaryrnd{CorpusFinishedKaggleCode} lines of R code.

@@

This section presents an overview of the usage of \eval in our corpus.
Richards et al. reported 82\% of JavaScript-enabled web pages using
\eval~\cite{ecoop11}. The case for R is more striking with 100\% of the
programs call the \eval function.\footnote{We combine calls to \c{eval},
  \c{eval.parent}, \c{evalq} and \c{local}.}  This high number is explained
by the fact that the R implementation, its core packages, use \eval
extensively. We differentiate uses based on their origin. A {\bf core} \eval
is one that occurs in the code of a core package.  A {\bf dependent} \eval
occurs in a package imported by the target package. An {\bf own} \eval is
one that occurs in the source code of the target package.

@@

%% SORRY  I HAD TO REFORMAT
\begin{table}[tb] \centering
 \begin{tabular}{r|c|c|c} \hline
Source & Core    & Dependent                      & Own \\\hline
CRAN   & 5.6M &  867K    & 259K \\
       & (1 / 337 / 46K)& (0 / 52 / 58K ) & (0 / 15 / 15K) \\\hline
Kaggle &  1.2M &  470K   & 0\\
 &  (0 / 721 / 18K) & (0 / 290 / 20K) & (0 / 0 / 0)\\\hline
\end{tabular}\caption{Eval calls in CRAN and Kaggle}\label{A}\end{table}

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. \OneCallSitesEnrich packages have a single call site, and one package, \emph{metafor}, has \MaxCallSitesEnrich
call sites of \eval. \NoCallSitesEnrich packages have no calls to \eval. The core packages
have \NbCoreEvalCallSites call sites of \eval and Kaggle has none. These numbers are
under-approximations as \eval may be called reflectively, through aliases,
or passed to a higher-order function.

@@

In this section we turn our attention to the expression passed for evaluation to
\eval. The \eval function accepts all types of R objects as inputs, but in the
majority of cases (\AllExpressionInputEvalCallPerc), it is called with
expressions. The remaining \AllValueInputEvalCallPerc are values, \ie, they
evaluate to themselves.
%
%Furthermore, the proportion of values passed to core
%\eval call is \missingTag{CoreEvalValueInputPerc} compared to a significantly higher
%\missingTag{PackageEvalValueInputPerc} for package \eval calls.
%
Passing a value into an eval is effectively an no-op so we look closely at the
expression types the call sites pass to eval across all the runs.
\AllValueInputEvalSitePerc of all call sites always pass values,
\AllExpressionInputEvalSitePerc pass expressions and a very large proportion,
\AllPolymorphicInputEvalSitePerc, pass both expressions and values as inputs
across all program runs. We inspected the eval callsites that accept only values
or both values and expression as inputs. The top three of these call sites
contribute to over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this function finds
the default values for parameters with those names in the caller function. Since
the default value for a function parameter can be a complex expression or a
benign value such as \c{NULL}, \c{match.arg} evaluates it through a call to
\eval. This makes the \eval inside \c{match.arg} ``polymorhpic''. The second
such \eval call originates from the popular \ggplot package of R which is a DSL
for plotting. \ggplot has it's own object system, \ggproto, which uses \eval
inside the \ggproto function to determine the super-class from which it
inherits. The default value of the superclass is \c{NULL} but in general could
be an arbitrarily complex expression. The third \eval call is in core R \c{str}
function. This function is used to print a summary of R objects. It internally
maps \eval function on the bindings of an environment which results in inputs of
multiples types being passed to \eval.

@@

We looked at the top ten expressions passed to core and package \eval calls. The
most frequent ten expressions to eval calls from core R contribute to 85\% of
all \eval calls.
\begin{table*}[tb]
  \centering
  \begin{tabular}{l|rr} \hline
    Expression & Eval Call & Eval Call \% \\\hline
    \c{c("auto", "shell", "radix")} & 1987105 & 29\%\\
    \c{c("auto", "shell", "quick", "radix")} & 1593169  & 23\%\\
    \c{\{info <- loadingNamespaceInfo(...} & 1008632 &       14\%\\
    \c{c("onLoad", "attach", "detach", "onUnload")}   & 470566 &      6.9\%\\
    \c{c("append", "prepend", "replace")} &              261587&       3.9\% \\
    \c{c("left", "right", "centre", "none")} & 162086     & 2.4\%\\
    \c{c("no", "ifany", "always")}   &                71580 &       1.1\%\\
    \c{c("pearson", "kendall", "spearman")}  & 72962 &      1.1\%\\
    \c{NULL}& 75330  &      1.1\% \\
    \c{Symbol}&                 66279&       1\%\\\hline
  \end{tabular}
  \caption{Top ten eval calls in Core}
  \label{A}
\end{table*}

@@

The most frequent ten expressions to eval calls from CRAN packages contribute to
77.1\% of all \eval calls.
\begin{table*}[tb]
  \centering
  \begin{tabular}{l|rr} \hline
    Expression & Eval Call & Eval Call \% \\\hline
    \c{Environment} &                                                              989302   & 61\%\\
    \c{column[rows] <<- what} &                                                    55677    & 3.5\%\\
    \c{function(value) freduce(value, `_function_list`)} &                         37251    & 2.3\%\\
    \c{NULL} &                         32005    & 2\%\\
    \c{List} &                         22293    & 1.4\%\\
    \c{c("default", "default2012", "default2011" ...}&                             20610    & 1.3\%\\
    \c{force(..1)}                                        &                        20461    & 1.3\%\\
    \c{alist(`_spec`)}                                   &                         18532    & 1.2\%\\
    \c{inner}                                           &                          18530    & 1.2\%\\
    \c{String Vector}                           &                           17487     & 1.1\%\\
  \end{tabular}
  \caption{Top ten eval calls in CRAN}
  \label{A}
\end{table*}

The expression \c{Environment} occurs because of the four callsites
explained above, \c{ggplot2::ggproto}, \c{R6::generator_funs},
\c{future::backtrace} and \c{RModel::str.RMmodel}. The next expression,
\c{column[rows] <<- what}, is used inside the \c{plyr::rbind.fill} function
to merge data frames by assigning concatenated vectors to rows. The \c{<<-}
operator is interesting in that it skips the current scope and assigns in a
parent scope in which the variable is already present. In our corpus, all
these \eval calls contribute to a single side-effect. The expression
\c{function(value) freduce(value, `_function_list`)} arises from the
\c{magrittr::\%>\%} function which is a pipe operator that pipes the output
of previous command to the next one. The expression is evaluated in a custom
environment to create a function binding for evaluating the components of
the pipe.\c{String Vector} and \c{List} also arise from the same function
when a string or a list is piped using the \c{\%>\%} function into the next
expression.  The \c{NULL} arises from \c{R6::generator_funs} function when
the \eval is passed a \c{NULL} argument by the \c{DataMask_generator}
package. The \c{c("default", "default2012", "default2011" ...} pattern
arises from \c{copula::polyG} where it reflectively access the default
expression for its formal parameter and evaluates it. The \c{force(..1)} and
\c{alist(`_spec`)} patterns occur in \c{glue::glue_data} function which
concatenates and interpolates strings. The two patterns occur because the
function captures unevaluated unnamed arguments and maps the evaluation of
\c{force(..1)} on them.  The \c{force} function forces
promises~\cite{oopsla19a} and returns the result of evaluation. The
\c{inner} pattern arises from \c{glue::identity_transformer} which enables
the creation of custom transformation functions for affecting the
interpolation and concatenation of input by the \c{glue} package.

\subsection{Operation Mix}
In this section, we look at the operations executed by the expressions and
symbols on evaluation. R is a lazy language, so symbols can be bound to
promises that can evaluate arbitrary code. All operations in R are function
calls, even primitive operations such as addition and assignment are treated
as calls to functions implemented in the interpreter. In our corpus, we
observe 2.9G calls to R functions of which 125M happen inside
\eval. Furthermore, unlike Javascript, R code also calls into peformant
native implementations of numerical and statistical routines. In the corpus
there are 463M calls to native code of which only 31M happens inside \eval.
Qualitatively, we observe all kinds of operations being executed inside
\eval calls, from numerical computing for statistical modeling to reflection
and metaprogramming for implementing DSLs. Even testing frameworks such
\c{genthat} heavily rely upon \eval to execute test code snippets. The core
R implementation is rife with uses of \eval for mundane things like
accessing default values of formal parameters and in statistical packages
for computation.

\subsubsection{Provenance}

In this section we turn our attention to the \AllExpressionInputEvalCallPerc
expressions passed as input to \eval. There are two ways to obtain
expressions in R.
\begin{itemize}
  \item \c{substitute} function can be used to synthesize ASTs from
    expressions by replacing symbols with their bindings in the specified
    environment.
  \item \c{expression} function can be used to create a vector of expression
    objects from text.
  \item \c{parse} function and its variants \c{str2expression} and
    \c{str2lang} can be used to turn strings or text from file into R
    expressions.
\end{itemize}

In our corpus, we observe 5.2\% cases where \eval directly evaluates the
output of \c{substitute}, 0.7\% cases where output of \c{parse} and its
variants is read directly and only 704 cases for \c{expression}. Majority of
expression objects consumed by \eval are generated by their transitive uses
by other functions.

The pattern \c{eval(parse(...))} can be used for dynamic code loading. This
forms the core of \c{source} and \c{sys.source} functions in R that are
commonly used for loading code in R files in interactive settings. We
investigated the number of cases in which the output of \c{parse} and its
variants is passed to \eval, directly or transitively by tainting their
output. This corresponds to \PercentParsedCallSites of the total \eval call
sites and \PercentParsedEvals of the eval calls. We observed that very few
of the eval call (\NbParseFilesRnd in total) consume the result of calling
\c{parse} on a file. Most of the eval calls consume the result of calling
\c{parse} on a string.  We also identified one function in core R,
\c{invokeRestartInteractively} that prompts the user for input, parses it,
and passes it to \eval.

@@

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \c{eval} can
access the local scope and the global scope. In R, \c{eval} has an argument
\c{envir} to indicate in which environment its first argument must be evaluated.
By default, it is equal to \c{parent.frame()}, 
which is the parent environment of \c{eval} call. This default argument
happens in \DefaultEnvirExprPercent of the eval calls.

This argument can also be \c{NULL}, a \c{List} or a data frame. This happens
in 3.5\% of the cases: \c{eval} copies the fields of the list or data frame
and creates bindings for them in a new environment. This pattern is used to
evaluate formulas which can directly refer to the fields of the data.  The
\c{envir} argument can also be a number $n$. It means that the environment
in which the expressions is evaluated will be the result of \c{sys.call(n)}
where $n$ refers to the $n$-th stack frame.

@@

The JavaScript paper identified a number of frequently occurring \eval
patterns which were detected using simple regular expressions on the \eval
strings. In the case of R, we match the patterns on the resolved AST.
